<!--
%\VignetteEngine{knitr::docco_classic}
%\VignetteIndexEntry{An Introduction to the animalmove package}
-->

Realized Mobility Index
=======================================


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra='style="display:block; margin: auto"', fig.align="center")
```

Load package library
```{r results='hide', message=FALSE, warning=FALSE}
library(animalmove)
```

Subsample Data
------------

Khulan dataset has been saved in the package data directory , and loaded on the package load.

Raw **khulan** dataset contains unlaltered original data.

```{r message=FALSE, warning=FALSE}
data(khulan.raw.data)
head(khulan.raw.data)
```

Number of rows in the khulan.raw.data
```{r message=FALSE, warning=FALSE}
nrow(khulan.raw.data)
```


For demonstration purposes we add a column that indicates a specie type to the data frame for each relocation data.

Enhanced dataset has been saved into khulan.test.data a column pop.type, which indicates a specie type.

Specie type assignment rule:  Individuals with ids 3,4,7, 6441 assigned to the specie1 type, and the rest is identified as specie2 type.

```{r message=FALSE, warning=FALSE}
data(khulan.test.data)
head(khulan.test.data)
```

Number of rows in the khulan.test.data

```{r message=FALSE, warning=FALSE}
nrow(khulan.test.data)
```

### Subsample data within time interval
--------------------------------------

We select at most 6 individuals within 2009, time interval 50 hours, and accuracy 50 hours, and subsampling scheme for Realized Mobility Index
```{r message=FALSE, warning=FALSE}
rmi.subsample.data <- subsample(khulan.test.data, start=c("2009-01-01 00:00"),end="2009-12-31 00:00",interval=c("50 hours"),accuracy=c("3 mins"),minIndiv=3,maxIndiv=6,mustIndiv=NULL,index.type="rmi")

head(rmi.subsample.data)
```

Note, a number of rows in the rmi subsample dataset

```{r message=FALSE, warning=FALSE}
nrow(rmi.subsample.data)

```

### Prepare data for the analysis in the spatial form
----------------------------

Create attribute data frame

```{r message=FALSE, warning=FALSE}
dt.rmi.data <- data.table(rmi.subsample.data)

dt.rmi.data.attr  <- dt.rmi.data[,list(id=id, pop.type=pop.type, x=X,y=Y, time=time, time.lag=time.lag, age=age, elevation= elevation, behaviour = behaviour)]
str(dt.rmi.data.attr)
```

Create spatial coordinates
```{r message=FALSE, warning=FALSE}
dt.rmi.data.xy <- dt.rmi.data[, list(x=X,y=Y)]

str(dt.rmi.data.xy)
```

Create spatial points
```{r message=FALSE, warning=FALSE}
xy.sp.rmi.data <- SpatialPoints(dt.rmi.data.xy)
```

Create spatial points data frame with attributes
```{r message=FALSE, warning=FALSE}
xy.rmi.data.spdf <- SpatialPointsDataFrame(xy.sp.rmi.data, dt.rmi.data.attr)
```

Create Individuals data.frame - relocations of khulan data
```{r message=FALSE, warning=FALSE}
khulan.reloc.spatial <- Individuals(xy.rmi.data.spdf, group.by="pop.type")
head(khulan.reloc.spatial)
```

Analysis
----------------------------

Compute individual mcp
```{r message=FALSE, warning=FALSE}
individual.mcp <- mcp(khulan.reloc.spatial[,1], percent = 100)
```

### Display individual mcp
----------------------------------

individual.mcp
```{r message=FALSE, warning=FALSE}
individual.mcp
```

### Compute population mcp
----------------------------------
```{r message=FALSE, warning=FALSE}
population.mcp <- mcp.population(khulan.reloc.spatial, percent = 100)
```

* Display population mcp
```{r message=FALSE, warning=FALSE}
population.mcp
```

### Compute Realized Mobility Index
----------------------------------
```{r message=FALSE, warning=FALSE}
rmi.index.population <- rmi.index(khulan.reloc.spatial, percent = 100, id = "id")

rmi.object <- RMIndex(rmi.index.population)
```

RMI Data 
```{r message=FALSE, warning=FALSE}
rmi.object@data
```

### RMI Summary
----------------------------------
```{r message=FALSE, warning=FALSE}
summary(rmi.object)
```
### Plot Results
----------------------------------

```{r message=FALSE, warning=FALSE}
par(mar = c(5,5, 2,0))
plot(rmi.object)
```

